---
title: "STRATAA Microbiome"
author: "Philip Ashton"
date: "`r Sys.Date()`"
output: 
  html_document:
      toc: true
      toc_float: true
      theme: united
      number_sections: true
---


```{r setup, error=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)

```

# STRATAA microbiome analysis

## Sources

The file handles are set in config.R as they're used by both this script and data_cleaning.

```{r source, error=FALSE}
source("/Users/flashton/Dropbox/GordonGroup/STRATAA_Microbiome/from_Leo/Leonardos_analysis/bin/00.core_functions.R")
source("/Users/flashton/Dropbox/GordonGroup/STRATAA_Microbiome/from_Leo/Leonardos_analysis/bin/config.R")
```

## Data import

```{r metadata_read_in, error=FALSE}
metadata <- read_metadata(metadata_handle)
# putting this here so that the output files of maaslin get named accroding to the variable names in the metadata file.
metadata <- metadata %>% mutate(Group = if_else(Group == 'Control_HealthySerosurvey', 'Household contact', Group)) %>% mutate(Group = if_else(Group == 'Acute_Typhi', 'Acute typhoid', Group)) %>% mutate(Group = if_else(Group == 'Carrier', 'High Vi-titre', Group))
```

## typhi vs healthy

### maaslin functional groups 

Functional groups associated with health

```{r read_in_maaslin_functional, error=FALSE}

bang_variables_for_analysis <- c("Group", "Sex", "Age", "Antibiotics_taken_before_sampling_assumptions")
mwi_variables_for_analysis <- c("Group", "Sex", "Age", "Antibiotics_taken_before_sampling_assumptions", "sequencing_lane")
# nep_variables_for_analysis <- c("Group", "Sex", "Age", "Antibiotics_taken_before_sampling_assumptions")
groups_to_analyse <- c('Acute_typhi', 'Control_HealthySerosurvey')

bang_func_maaslin <- read_in_maaslin('Bangladesh', groups_to_analyse, bang_variables_for_analysis, 'bigmap')
mal_func_maaslin <- read_in_maaslin('Malawi', groups_to_analyse, mwi_variables_for_analysis, 'bigmap')
nep_func_maaslin <- read_in_maaslin('Nepal', groups_to_analyse, bang_variables_for_analysis, 'bigmap')
```

Just Malawi and Bangladesh.

```{r maaslin_functional_groups, error=FALSE}
##
bang_mal <- list(bang_func_maaslin, mal_func_maaslin)
combined_results <- run_inner_join_maaslins(bang_mal, c('_bang', '_mal'), mwi_variables_for_analysis, groups_to_analyse, 'bigmap', maaslin_functional_output_root_folder)

# View(combined_results$positive_coef)
positive_table <- combined_results$positive_coef %>%
  dplyr::select(-metadata, -value, -N_bang, -N.not.0_bang, -pval_bang, -N_mal, -N.not.0_mal, -pval_mal) %>% 
  rename(`Coefficient Bangladesh` = coef_bang, `Standard Error Bangladesh` = stderr_bang, `Q-value Bangladesh` = qval_bang, 
         `Coefficient Malawi` = coef_mal, `Standard Error Malawi` = stderr_mal, `Q-value Malawi` = qval_mal) %>% 
  dplyr::mutate_if(is.numeric, ~as.character(signif(., 3))) %>% 
  kbl() %>% 
  kable_styling()

# Process negative coefficients
negative_table <- combined_results$negative_coef %>%
  dplyr::select(-metadata, -value, -N_bang, -N.not.0_bang, -pval_bang, -N_mal, -N.not.0_mal, -pval_mal) %>% 
  rename(`Coefficient Bangladesh` = coef_bang, `Standard Error Bangladesh` = stderr_bang, `Q-value Bangladesh` = qval_bang, 
         `Coefficient Malawi` = coef_mal, `Standard Error Malawi` = stderr_mal, `Q-value Malawi` = qval_mal) %>% 
  dplyr::mutate_if(is.numeric, ~as.character(signif(., 3))) %>% 
  kbl() %>% 
  kable_styling()

# Split species and clean data for positive coefficients
positive_cleaned <- combined_results$positive_coef %>% 
  mutate(split_species = str_split(Species, '_')) %>% 
  mutate(genus = sapply(split_species, "[[", 1)) %>% 
  mutate(specific = sapply(split_species, "[[", 2)) %>% 
  mutate(clean_species = paste(genus, specific, sep = ' ')) %>% 
  dplyr::select(-split_species, -genus, -specific) %>% 
  relocate(clean_species, .after = Species) %>% 
  group_by(MGC_class) %>% 
  arrange(Species) %>% 
  summarise(n = n(), Species = paste(clean_species, collapse = ', ')) %>% 
  arrange(desc(n), MGC_class) %>%  
  kbl() %>% 
  kable_styling()

# Split species and clean data for negative coefficients
negative_cleaned <- combined_results$negative_coef %>% 
  mutate(split_species = str_split(Species, '_')) %>% 
  mutate(genus = sapply(split_species, "[[", 1)) %>% 
  mutate(specific = sapply(split_species, "[[", 2)) %>% 
  mutate(clean_species = paste(genus, specific, sep = ' ')) %>% 
  dplyr::select(-split_species, -genus, -specific) %>% 
  relocate(clean_species, .after = Species) %>% 
  group_by(MGC_class) %>% 
  arrange(Species) %>% 
  summarise(n = n(), Species = paste(clean_species, collapse = ', ')) %>% 
  arrange(desc(n), MGC_class) %>%  
  kbl() %>% 
  kable_styling()

# Display the tables
positive_table
negative_table
positive_cleaned
negative_cleaned

```

There were `r nrow(combined_results$positive_coef)` MGCs significantly (q-val < 0.05) associated with health/disease in Malawi and Bangladesh.

```{r}

```

## healthy vs carrier



### read in data
```{r read_in_maaslin_functional_healthy_carrier, error=FALSE}
# bang_variables_for_analysis <- c("Group", "Sex", "Age")
mwi_variables_for_analysis <- c("Group", "Sex", "Age", "sequencing_lane")
# nep_variables_for_analysis <- c("Group", "Sex", "Age", "Antibiotics_taken_before_sampling_assumptions")
groups_to_analyse <- c('High Vi-titre', 'Household contact')

# bang_func_carr_health_maaslin <- read_in_maaslin('Bangladesh', groups_to_analyse, bang_variables_for_analysis, 'bigmap')
mal_func_carr_health_maaslin <- read_in_maaslin('Malawi', groups_to_analyse, mwi_variables_for_analysis, 'bigmap')
```

### maaslin
```{r maaslin_functional_groups_healthy_carrier, error=FALSE}
# combined_results <- run_inner_join_maaslins(bang_func_carr_health_maaslin, mal_func_carr_health_maaslin, bang_variables_for_analysis, mwi_variables_for_analysis, groups_to_analyse, 'bigmap', maaslin_functional_output_root_folder)

# bang_mal <- list(bang_func_carr_health_maaslin, mal_func_carr_health_maaslin)
# combined_results <- run_inner_join_maaslins(bang_mal, c('_bang', '_mal'), mwi_variables_for_analysis, groups_to_analyse, 'bigmap', maaslin_functional_output_root_folder)

# # positive is associated with health
# # combined_results$positive_coef %>%  kbl() %>% kable_styling()
# combined_results$positive_coef %>% 
#   select(!c(metadata, value, N_bang, N.not.0_bang, pval_bang, N_mal, N.not.0_mal, pval_mal)) %>% 
#   rename(`Coefficient Bangladesh` = coef_bang, `Standard Error Bangladesh` = stderr_bang, `Q-value Bangladesh` = qval_bang, `Coefficient Malawi` = coef_mal, `Standard Error Malawi` = stderr_mal, `Q-value Malawi` = qval_mal) %>% 
#   dplyr::mutate_if(is.numeric, funs(as.character(signif(., 3)))) %>% 
#   kbl() %>% kable_styling()
# # negative is associated with carrier
# # combined_results$negative_coef %>%  kbl() %>% kable_styling()
# combined_results$negative_coef %>% 
#   select(!c(metadata, value, N_bang, N.not.0_bang, pval_bang, N_mal, N.not.0_mal, pval_mal)) %>% 
#   rename(`Coefficient Bangladesh` = coef_bang, `Standard Error Bangladesh` = stderr_bang, `Q-value Bangladesh` = qval_bang, `Coefficient Malawi` = coef_mal, `Standard Error Malawi` = stderr_mal, `Q-value Malawi` = qval_mal) %>% 
#   dplyr::mutate_if(is.numeric, funs(as.character(signif(., 3)))) %>% 
#   kbl() %>% kable_styling()


# # print out a nicely formatted table of the count of the MGCs
# combined_results$positive_coef %>% mutate(split_species = str_split(Species, '_')) %>% mutate(genus = sapply(split_species, "[[", 1)) %>% mutate(specific = sapply(split_species, "[[", 2)) %>% mutate(clean_species = paste(genus, specific, sep = ' ')) %>% select(-split_species, -genus, -specific) %>% relocate(clean_species, .after = Species) %>% group_by(MGC_class) %>% arrange(Species) %>% summarise(n = n(), Species = paste(clean_species, collapse = ', ')) %>% arrange(desc(n), MGC_class) %>%  kbl() %>% kable_styling()
# combined_results$negative_coef %>% mutate(split_species = str_split(Species, '_')) %>% mutate(genus = sapply(split_species, "[[", 1)) %>% mutate(specific = sapply(split_species, "[[", 2)) %>% mutate(clean_species = paste(genus, specific, sep = ' ')) %>% select(-split_species, -genus, -specific) %>% relocate(clean_species, .after = Species) %>% group_by(MGC_class) %>% arrange(Species) %>% summarise(n = n(), Species = paste(clean_species, collapse = ', ')) %>% arrange(desc(n), MGC_class) %>%  kbl() %>% kable_styling()
```