---
title: "STRATAA Microbiome"
author: "Philip Ashton"
date: "`r Sys.Date()`"
output: 
  html_document:
      toc: true
      toc_float: true
      theme: united
      number_sections: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE)

```

# STRATAA microbiome analysis

## Imports
```{r}
library(rbiom)
library(kableExtra)
library(ggplot2)
library(ggpubr)
library(patchwork)
library(vegan)
```


## Sources

The file handles are set in config.R as they're used by both this script and data_cleaning.

```{r source}
source("/Users/flashton/Dropbox/GordonGroup/STRATAA_Microbiome/from_Leo/Leonardos_analysis/bin/core_functions.R")
source("/Users/flashton/Dropbox/GordonGroup/STRATAA_Microbiome/from_Leo/Leonardos_analysis/bin/config.R")
```


## Metadata basics

First do the plots with kruskall wallis comparison between groups to see if there's an overall difference, then do the pairwise tests because KW says there is a difference.


```{r metadata_basics}
metadata <- read_metadata(metadata_handle)

metadata <- metadata %>% mutate(Group = if_else(Group == 'Control_HealthySerosurvey', 'Healthy Control', Group))
metadata <- metadata %>% mutate(Group = if_else(Group == 'Acute_Typhi', 'Acute typhoid', Group))
number_per_country <- metadata %>% group_by(Country) %>% summarise(count = n())

number_per_country %>% kbl() %>% kable_styling()

print(sum(number_per_country$count))

metadata %>% group_by(Group) %>% summarise(count = n()) %>% kbl() %>% kable_styling()

baseline_chars <- get_baseline_characteristics(metadata)

baseline_chars$pct_anti
baseline_chars$pct_anti %>% na_if(0)


# BEWARE - na_if(0.0) will convert ALL 0s to NAs, this is ok as they're only in the antibiotics of carriers and controls
# at the moment, but if others get added in, will screw with it.
# baseline_chars %>% rename(`Median age` = median_age, `Women (%)` = pct_fem, `Antibiotics in last 2 weeks (%)` = pct_anti) %>% pivot_longer(!c(Country, Group)) %>% rename(variable_name = name) %>% pivot_wider(names_from = Group, values_from = value) %>% filter(variable_name != 'number') %>% na_if(0.0) %>% kbl(digits = c(NA, NA, 1, 1, 1)) %>% kable_styling()

# na_if stopped working, so had to do this instead
baseline_chars_table <- baseline_chars %>% rename(`Median age` = median_age, `Women (%)` = pct_fem, `Antibiotics in last 2 weeks (%)` = pct_anti) %>% pivot_longer(!c(Country, Group)) %>% rename(variable_name = name) %>% pivot_wider(names_from = Group, values_from = value) %>% filter(variable_name != 'number') 

baseline_chars_table$Carrier <- replace(baseline_chars_table$Carrier,which(baseline_chars_table$Carrier==0),NA)
baseline_chars_table$`Healthy Control` <- replace(baseline_chars_table$`Healthy Control`,which(baseline_chars_table$`Healthy Control`==0),NA)
baseline_chars_table %>% kbl(digits = c(NA, NA, 1, 1, 1)) %>% kable_styling()

ggplot(metadata, aes(x = Country, y = Age, fill = Group)) + geom_boxplot() + stat_compare_means(method = 'kruskal.test', label = "p")
comparisons_groups <- list(c("Acute typhoid", "Carrier"), c("Acute typhoid", "Healthy Control"), c("Carrier", "Healthy Control"))
comparisons_countries <- list(c('Bangladesh', 'Malawi'), c('Bangladesh', 'Nepal'), c('Malawi', 'Nepal'))
# default stat method when doing pairwise tests in wilcoxon.
ggboxplot(metadata, facet.by = "Country", y = "Age", x = "Group", color = "Group") + stat_compare_means(comparisons = comparisons_groups) + rremove("x.text") + rremove("xlab") + rremove("x.ticks") # +rotate_x_text(angle = 45)
ggboxplot(metadata, facet.by = "Group", y = "Age", x = "Country", color = "Country") + stat_compare_means(comparisons = comparisons_countries) + rotate_x_text(angle = 45)

```

```{r metadata_basic_plots, fig.width = 4, fig.height= 8}
country_group_sex <- metadata %>% group_by(Country, Group, Sex) %>% summarise(count = n()) 

plot_sex <- function(eg1, c){
  d <- eg1 %>% filter(Country == c)
  p <- ggplot(d, aes(x = Group, y = count, fill = Sex)) + 
    geom_bar(stat ='identity', position = 'fill') + 
    ylab('Proportion') + 
    ggtitle(c)# +
    #theme(axis.text=element_text(size=34), axis.title=element_text(size=36,face="bold"), plot.title = element_text(size = 40, face = "bold"), legend.key.size = unit(4, 'cm'), legend.title = element_text(size = 34), legend.text = element_text(size = 28))
  
  return(p)
}

m_sex <- plot_sex(country_group_sex, 'Malawi')
b_sex <- plot_sex(country_group_sex, 'Bangladesh')
n_sex <- plot_sex(country_group_sex, 'Nepal')
m_sex / b_sex / n_sex 
```



## Alpha diversity - metaphlan

```{r read_in_beta_diversity}
strataa_metaphlan_data <- read.csv(file = file.path(metaphlan_input_folder, '2023.05.11.all_strataa_metaphlan.csv'), header= TRUE, sep = ",", row.names = 1, stringsAsFactors = FALSE, check.names=FALSE)
strataa_metaphlan_metadata <- read.csv(file = file.path(metaphlan_input_folder, '2023.05.11.strataa_metadata.metaphlan.csv'), header = TRUE, sep = ",", row.names = 1, stringsAsFactors = FALSE)
strataa_metaphlan_metadata <- strataa_metaphlan_metadata %>% mutate(SampleID = rownames(strataa_metaphlan_metadata))
```
Stats
```{r alpha_diversity}

all_three_all_groups_alpha <- metaphlan_alpha(strataa_metaphlan_data, strataa_metaphlan_metadata, countries_of_interest = c('Bangladesh', 'Malawi', 'Nepal'), groups_of_interest = c('Acute_Typhi', 'Carrier', 'Control_HealthySerosurvey'), comparisons = list(c('Bangladesh', 'Malawi'), c('Malawi', 'Nepal'), c('Bangladesh', 'Nepal')))


```
Plots
```{r alpha_diversity}

country_comps <- list(c('Bangladesh', 'Malawi'), c('Malawi', 'Nepal'), c('Bangladesh', 'Nepal'))
ggplot(all_three_shannon_alpha, aes(x=Country, y=alpha)) + 
  ggtitle("By country") + 
  labs(x="", y="Alpha diversity") +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12), axis.text.y = element_text(size = 12), axis.title = element_text(size = 14, face = 'bold')) +
  stat_compare_means(size = 4, label = "p.format", comparisons = country_comps)

group_comps <- list(c(1, 2), c(2, 3), c(1, 3))
ggplot(all_three_shannon_alpha, aes(x=Group, y=alpha)) + 
  ggtitle("By group") + 
  labs(x="", y="Alpha diversity") +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12), axis.text.y = element_text(size = 12), axis.title = element_text(size = 14, face = 'bold')) +
  stat_compare_means(size = 4, label = "p.format", comparisons = group_comps)
  #geom_point(alpha = 0.3, position = "jitter")
```

## Alpha diversity

```{r alpha_diversity}
meta_for_alpha <- metadata %>% select(isolate, Group, Sex, Country, Antibiotics_taken_before_sampling_yes_no_assumptions, age_bracket)

```


### all three

ANOVA results

```{r alpha_anova_read_in}
anova_all_three_csgaa_res <- read_delim(file.path(output_folder_all_three, '3_alpha', 'shannon.anova.results.tsv'), delim = '\t')
anova_all_three_csgaa_res %>% kbl %>% kable_styling()
```

Plots

```{r alpha_anova}
all_three_shannon_alpha <- read_delim(file.path(output_folder_all_three, '3_alpha', 'shannon.alpha_results.tsv'))
all_three_shannon_alpha <- left_join(all_three_shannon_alpha, meta_for_alpha, by = 'isolate', )
all_three_shannon_alpha <- all_three_shannon_alpha %>% mutate(Group = if_else(Group == 'Control_HealthySerosurvey', 'Healthy Control', Group))
all_three_shannon_alpha <- all_three_shannon_alpha %>% mutate(Group = if_else(Group == 'Acute_Typhi', 'Acute typhoid', Group)) %>% arrange(Country)


country_comps <- list(c('Bangladesh', 'Malawi'), c('Malawi', 'Nepal'), c('Bangladesh', 'Nepal'))
ggplot(all_three_shannon_alpha, aes(x=Country, y=alpha)) + 
  ggtitle("By country") + 
  labs(x="", y="Alpha diversity") +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12), axis.text.y = element_text(size = 12), axis.title = element_text(size = 14, face = 'bold')) +
  stat_compare_means(size = 4, label = "p.format", comparisons = country_comps)

group_comps <- list(c(1, 2), c(2, 3), c(1, 3))
ggplot(all_three_shannon_alpha, aes(x=Group, y=alpha)) + 
  ggtitle("By group") + 
  labs(x="", y="Alpha diversity") +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12), axis.text.y = element_text(size = 12), axis.title = element_text(size = 14, face = 'bold')) +
  stat_compare_means(size = 4, label = "p.format", comparisons = group_comps)
  #geom_point(alpha = 0.3, position = "jitter")

country_group_comps <- list(c('Acute typhoid', 'Carrier'), c('Acute typhoid', 'Healthy Control'), c('Healthy Control', 'Carrier'))

ggboxplot(all_three_shannon_alpha, facet.by = "Country", y = "alpha", x = "Group", color = "Group") + stat_compare_means(comparisons = country_group_comps, label = 'p.signif') + rremove("x.text") + rremove("xlab") + rremove("x.ticks") # +rotate_x_text(angle = 45)


ggboxplot(all_three_shannon_alpha, facet.by = "Group", y = "alpha", x = "Country", color = "Country") + stat_compare_means(comparisons = country_comps) + rremove("x.text") + rremove("xlab") + rremove("x.ticks") # +rotate_x_text(angle = 45)


antibiotic_comps <- list(c('Yes', 'No'))
ggboxplot(all_three_shannon_alpha, facet.by = "Country", y = "alpha", x = "Antibiotics_taken_before_sampling_yes_no_assumptions", color = "Antibiotics_taken_before_sampling_yes_no_assumptions", legend.title = 'Antibiotics') + stat_compare_means(comparisons = antibiotic_comps) + rremove("x.text") + rremove("xlab") + rremove("x.ticks")


```

### blantyre

ANOVA

```{r blantre_alpha_anova}

anova_blantyre_sgaa_res <- read_delim(file.path(output_folder_blantyre, '3_alpha', 'shannon.anova.results.tsv'), delim = '\t')
anova_blantyre_sgaa_res %>% kbl %>% kable_styling()

```

Plots

- The p-value on the graph is a kruskal wallis test (which is just wilcoxon i think?)

```{r alpha_blantyre_plots}
blantyre_shannon_alpha <- read_delim(file.path(output_folder_blantyre, '3_alpha', 'shannon.alpha_results.tsv'))
blantyre_shannon_alpha <- left_join(blantyre_shannon_alpha, meta_for_alpha, by = 'isolate')

my_comparisons <- list(c('No', 'Yes'), c('No', 'Yes'), c('No', 'Yes'))

ggplot(blantyre_shannon_alpha, aes(x=age_bracket, y=alpha, fill = Antibiotics_taken_before_sampling_yes_no_assumptions)) + 
  #ggtitle("All three countries - country by group") + 
  labs(x="Age bracket", y="Alpha diversity", fill='Antibiotics', color = 'Antibiotics') +
  geom_boxplot(outlier.shape = NA, alpha = 0.2) +
  #geom_point(aes(color=as.factor(Antibiotics_taken_before_sampling_yes_no_assumptions)), alpha = 0.3) +
  geom_point(alpha = 0.5, position=position_dodge(width=0.75),aes(group=Antibiotics_taken_before_sampling_yes_no_assumptions, color=as.factor(Antibiotics_taken_before_sampling_yes_no_assumptions))) +
  stat_compare_means(label.y = c(5.2, 5.5, 5.7), size = 4, label = "p.format")
  

```


### dhaka

```{r}

anova_dhaka_sgaa_res <- read_delim(file.path(output_folder_dhaka, '3_alpha', 'shannon.anova.results.tsv'), delim = '\t')
anova_dhaka_sgaa_res %>% kbl %>% kable_styling()

```

### kathmandu

```{r}

anova_kathmandu_sgaa_res <- read_delim(file.path(output_folder_kathmandu, '3_alpha', 'shannon.anova.results.tsv'), delim = '\t')
anova_kathmandu_sgaa_res %>% kbl %>% kable_styling()

```





## Beta diversity - metaphlan

All groups.
```{r}
all_countries_beta_all_groups <- metaphlan_beta(strataa_metaphlan_data, strataa_metaphlan_metadata, c('Bangladesh', 'Malawi', 'Nepal'), c('Acute_Typhi', 'Carrier', 'Control_HealthySerosurvey'))
all_countries_beta_all_groups$pn_res %>% kbl %>% kable_styling()

bgd_beta_all_groups <- metaphlan_beta(strataa_metaphlan_data, strataa_metaphlan_metadata, c('Bangladesh'), c('Acute_Typhi', 'Carrier', 'Control_HealthySerosurvey'))
bgd_beta_all_groups$pn_res %>% kbl %>% kable_styling()
# bgd_beta_all_groups$pc12
# bgd_beta_all_groups$pc34

mal_beta_all_groups <- metaphlan_beta(strataa_metaphlan_data, strataa_metaphlan_metadata, c('Malawi'), c('Acute_Typhi', 'Carrier', 'Control_HealthySerosurvey'))
mal_beta_all_groups$pn_res %>% kbl %>% kable_styling()
# mal_beta_all_groups$pc12
# mal_beta_all_groups$pc34



nep_beta_all_groups <- metaphlan_beta(strataa_metaphlan_data, strataa_metaphlan_metadata, c('Nepal'), c('Acute_Typhi', 'Carrier', 'Control_HealthySerosurvey'))
nep_beta_all_groups$pn_res %>% kbl %>% kable_styling()


all_countries_beta_all_groups$pc12 / bgd_beta_all_groups$pc12 / nep_beta_all_groups$pc12

```

Acute vs healthy.
```{r}

all_countries_beta_acute_healthy <- metaphlan_beta(strataa_metaphlan_data, strataa_metaphlan_metadata, c('Bangladesh', 'Malawi', 'Nepal'), c('Acute_Typhi', 'Control_HealthySerosurvey'))
all_countries_beta_acute_healthy$pn_res %>% kbl %>% kable_styling()

bgd_beta_acute_healthy <- metaphlan_beta(strataa_metaphlan_data, strataa_metaphlan_metadata, c('Bangladesh'), c('Acute_Typhi', 'Control_HealthySerosurvey'))
bgd_beta_acute_healthy$pn_res %>% kbl %>% kable_styling()

mal_beta_acute_healthy <- metaphlan_beta(strataa_metaphlan_data, strataa_metaphlan_metadata, c('Malawi'), c('Acute_Typhi', 'Control_HealthySerosurvey'))
mal_beta_acute_healthy$pn_res %>% kbl %>% kable_styling()

nep_beta_acute_healthy <- metaphlan_beta(strataa_metaphlan_data, strataa_metaphlan_metadata, c('Nepal'), c('Acute_Typhi', 'Control_HealthySerosurvey'))
nep_beta_acute_healthy$pn_res %>% kbl %>% kable_styling()

all_countries_beta_acute_healthy$pc12 + bgd_beta_acute_healthy$pc12 + mal_beta_acute_healthy$pc12 + nep_beta_acute_healthy$pc12
```





It looks like some might be being trimmed by the coordinate limits in the plot_beta function, but they're not (i've checked both blantyre and dhaka).

```{r fig.height= 3, fig.width = 3, eval=FALSE}
#b_g / d_g | k_g / plot_spacer() + plot_layout(guides = 'collect')
# b_g + d_g + k_g + guide_area() + plot_layout(ncol = 2, guides = 'collect')
```


## Taxa associated with participant groups

### EdgeR

```{r, eval = FALSE}
combined_output_root <- '/Users/flashton/Dropbox/GordonGroup/STRATAA_Microbiome/from_leo/phil/2022.11.08/combined_analysis'
if (!dir.exists(combined_output_root)){ dir.create(combined_output_root) }


bangladesh_covars <- c('Country', 'Sex', 'Age', 'Antibiotics_taken_before_sampling_yes_no_assumptions')
malawi_covars <- c('Country', 'Sex', 'Age', 'Antibiotics_taken_before_sampling_yes_no_assumptions', 'sequencing_run')

run_combine_edgeR(groups_for_comparison = c("Control_HealthySerosurvey", "Acute_Typhi"), bangladesh_covars, malawi_covars)
run_combine_edgeR(groups_for_comparison = c("Control_HealthySerosurvey", "Carrier"), bangladesh_covars, malawi_covars)

```

### maaslin taxonomic

Combine the taxonomic maaslins, and print out the species that are sig in both and share directions.

Because sequencing run and participant type are totally confounded for Bangladesh, need to exclude sequencing run from the final model for Bangladesh (otherwise, wipes out the signals).

between acute <-> health.

```{r, eval = FALSE}
groups_to_analyse <- c('Acute_Typhi', 'Control_HealthySerosurvey')
mwi_variables_for_analysis <- c("Group", "Sex", "Age", "Antibiotics_taken_before_sampling_yes_no_assumptions", "sequencing_lane")
bang_variables_for_analysis <- c("Group", "Sex", "Age", "Antibiotics_taken_before_sampling_yes_no_assumptions")
run_combine_maaslins(groups_to_analyse, mwi_variables_for_analysis, bang_variables_for_analysis, maaslin_taxonomic_output_root_folder, 'metaphlan')
```


```{r, eval = FALSE}
groups_to_analyse <- c('Carrier', 'Control_HealthySerosurvey')
mwi_variables_for_analysis <- c("Group", "Sex", "Age", "sequencing_lane")
bang_variables_for_analysis <- c("Group", "Sex", "Age")
run_combine_maaslins(groups_to_analyse, mwi_variables_for_analysis, bang_variables_for_analysis, maaslin_taxonomic_output_root_folder, 'metaphlan')
```


Print the results for all the P. copri clades

```{r, eval = FALSE}
all_taxa_maaslin <- combined_maaslins$all_features
p_copri_maaslin <- all_taxa_maaslin %>% filter(grepl('_Prevotella_copri_', feature)) %>% filter(qval_bang < 0.05 | qval_mal < 0.05)
write_csv(p_copri_maaslin, file.path(maaslin_taxonomic_output_root_folder, 'combined_maaslins_p_copri.csv'))
```

## maaslin functional groups 

Functional groups associated with health


```{r, eval = FALSE}
# This needs to be replaced by combine_maaslins

variables_for_analysis <- c("Group", "Gender", "Age", "Antibiotics_taken_before_sampling_yes_no_assumptions")
vars_for_dirname <- paste(variables_for_analysis, collapse = '.')
groups_to_analyse <- c('Acute_Typhi', 'Control_Healthy')

bang_maaslin_output_dir <- file.path(maaslin_functional_output_root_folder, paste('Bangladesh', paste(groups_to_analyse, collapse = '_vs_'), vars_for_dirname, sep = '_'))

malawi_maaslin_output_dir <- file.path(maaslin_functional_output_root_folder, paste('Malawi', paste(groups_to_analyse, collapse = '_vs_'), vars_for_dirname, sep = '_'))

bang_maaslin <- read_delim(file.path(bang_maaslin_output_dir, "all_results.tsv"), delim = "\t", escape_double = FALSE, trim_ws = TRUE)
malawi_maaslin <- read_delim(file.path(malawi_maaslin_output_dir, "all_results.tsv"), delim = "\t", escape_double = FALSE, trim_ws = TRUE)

combined_maaslins <- combine_maaslins(bang_maaslin, malawi_maaslin)

combined_maaslins_positive_coef <- combined_maaslins$positive_coef
combined_maaslins_negative_coef <- combined_maaslins$negative_coef

# hacky as hell thing to split the feature name to extract the organism and the enzyme class.
# because separate isn't working with .. or \\.\\. which is the actual separator used.
combined_maaslins_positive_coef <- combined_maaslins_positive_coef %>% separate_wider_delim(feature, delim = 'Entryname.', names_sep = '', cols_remove = FALSE) %>% separate_wider_delim(feature2, delim = '..OS.', names_sep = '') %>% separate_wider_delim(feature22, delim = '..SMASH', names_sep = '') %>% select(!c(feature1, feature222)) %>% rename(MGC_class = feature21, Species = feature221, feature = featurefeature)
combined_maaslins_negative_coef <- combined_maaslins_negative_coef %>% separate_wider_delim(feature, delim = 'Entryname.', names_sep = '', cols_remove = FALSE) %>% separate_wider_delim(feature2, delim = '..OS.', names_sep = '') %>% separate_wider_delim(feature22, delim = '..SMASH', names_sep = '') %>% select(!c(feature1, feature222)) %>% rename(MGC_class = feature21, Species = feature221, feature = featurefeature)

combined_maaslins_positive_coef %>%  kbl() %>% kable_styling()
combined_maaslins_negative_coef %>%  kbl() %>% kable_styling()

write_csv(combined_maaslins_positive_coef, file.path(maaslin_functional_output_root_folder, 'combined_maaslins_positive_coef.csv'))
combined_maaslins_negative_coef %>%  kbl() %>% kable_styling()
write_csv(combined_maaslins_negative_coef, file.path(maaslin_functional_output_root_folder, 'combined_maaslins_negative_coef.csv'))

```

```{r, eval = FALSE}

groups_to_analyse <- c('Carrier', 'Control_HealthySerosurvey')
mwi_variables_for_analysis <- c("Group", "Sex", "Age")
bang_variables_for_analysis <- c("Group", "Sex", "Age")
run_combine_maaslins(groups_to_analyse, mwi_variables_for_analysis, bang_variables_for_analysis, maaslin_functional_output_root_folder, 'bigmap')
```

