---
title: "STRATAA Microbiome"
author: "Philip Ashton"
date: "`r Sys.Date()`"
output: 
  html_document:
      toc: true
      toc_float: true
      theme: united
      number_sections: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE, warning=FALSE, cache = TRUE)

```

# STRATAA microbiome analysis

## Imports
```{r}
library(kableExtra)
library(ggplot2)
library(ggpubr)
library(patchwork)
library(vegan)
```


## Sources

The file handles are set in config.R as they're used by both this script and data_cleaning.

```{r}
source("/Users/flashton/Dropbox/GordonGroup/STRATAA_Microbiome/from_Leo/Leonardos_analysis/bin/core_functions.R")
source("/Users/flashton/Dropbox/GordonGroup/STRATAA_Microbiome/from_Leo/Leonardos_analysis/bin/config.R")
```


## Metadata basics

First do the plots with kruskall wallis comparison between groups to see if there's an overall difference, then do the pairwise tests because KW says there is a difference.


```{r}
metadata <- read_metadata(metadata_handle)

metadata <- metadata %>% mutate(Group = if_else(Group == 'Control_HealthySerosurvey', 'Healthy Control', Group))
metadata <- metadata %>% mutate(Group = if_else(Group == 'Acute_Typhi', 'Acute typhoid', Group))
number_per_country <- metadata %>% group_by(Country) %>% summarise(count = n())

number_per_country %>% kbl() %>% kable_styling()

print(sum(number_per_country$count))

metadata %>% group_by(Group) %>% summarise(count = n()) %>% kbl() %>% kable_styling()

baseline_chars <- get_baseline_characteristics(metadata)

baseline_chars$pct_anti
baseline_chars$pct_anti %>% na_if(0)

# BEWARE - na_if(0.0) will convert ALL 0s to NAs, this is ok as they're only in the antibiotics of carriers and controls
# at the moment, but if others get added in, will screw with it.
baseline_chars %>% rename(`Median age` = median_age, `Women (%)` = pct_fem, `Antibiotics in last 2 weeks (%)` = pct_anti) %>% pivot_longer(!c(Country, Group)) %>% rename(variable_name = name) %>% pivot_wider(names_from = Group, values_from = value) %>% filter(variable_name != 'number') %>% na_if(0.0) %>% kbl(digits = c(NA, NA, 1, 1, 1)) %>% kable_styling()


ggplot(metadata, aes(x = Country, y = Age, fill = Group)) + geom_boxplot() + stat_compare_means(method = 'kruskal.test', label = "p")
comparisons_groups <- list(c("Acute typhoid", "Carrier"), c("Acute typhoid", "Healthy Control"), c("Carrier", "Healthy Control"))
comparisons_countries <- list(c('Bangladesh', 'Malawi'), c('Bangladesh', 'Nepal'), c('Malawi', 'Nepal'))
# default stat method when doing pairwise tests in wilcoxon.
ggboxplot(metadata, facet.by = "Country", y = "Age", x = "Group", color = "Group") + stat_compare_means(comparisons = comparisons_groups) + rremove("x.text") + rremove("xlab") + rremove("x.ticks") # +rotate_x_text(angle = 45)
ggboxplot(metadata, facet.by = "Group", y = "Age", x = "Country", color = "Country") + stat_compare_means(comparisons = comparisons_countries) + rotate_x_text(angle = 45)



```

```{r fig.width = 4, fig.height= 8}
country_group_sex <- metadata %>% group_by(Country, Group, Sex) %>% summarise(count = n()) 

plot_sex <- function(eg1, c){
  d <- eg1 %>% filter(Country == c)
  p <- ggplot(d, aes(x = Group, y = count, fill = Sex)) + 
    geom_bar(stat ='identity', position = 'fill') + 
    ylab('Proportion') + 
    ggtitle(c)# +
    #theme(axis.text=element_text(size=34), axis.title=element_text(size=36,face="bold"), plot.title = element_text(size = 40, face = "bold"), legend.key.size = unit(4, 'cm'), legend.title = element_text(size = 34), legend.text = element_text(size = 28))
  
  return(p)
}

m_sex <- plot_sex(country_group_sex, 'Malawi')
b_sex <- plot_sex(country_group_sex, 'Bangladesh')
n_sex <- plot_sex(country_group_sex, 'Nepal')
m_sex / b_sex / n_sex 
```



## Alpha diversity

```{r}
meta_for_alpha <- metadata %>% select(isolate, Group, Sex, Country, Antibiotics_taken_before_sampling_yes_no_assumptions, age_bracket)

```


### all three

ANOVA results

```{r}
anova_all_three_csgaa_res <- read_delim(file.path(output_folder_all_three, '3_alpha', 'shannon.anova.results.tsv'), delim = '\t')
anova_all_three_csgaa_res %>% kbl %>% kable_styling()
```

Plots

```{r}
all_three_shannon_alpha <- read_delim(file.path(output_folder_all_three, '3_alpha', 'shannon.alpha_results.tsv'))
all_three_shannon_alpha <- left_join(all_three_shannon_alpha, meta_for_alpha, by = 'isolate', )
all_three_shannon_alpha <- all_three_shannon_alpha %>% mutate(Group = if_else(Group == 'Control_HealthySerosurvey', 'Healthy Control', Group))
all_three_shannon_alpha <- all_three_shannon_alpha %>% mutate(Group = if_else(Group == 'Acute_Typhi', 'Acute typhoid', Group)) %>% arrange(Country)


country_comps <- list(c('Bangladesh', 'Malawi'), c('Malawi', 'Nepal'), c('Bangladesh', 'Nepal'))
ggplot(all_three_shannon_alpha, aes(x=Country, y=alpha)) + 
  ggtitle("By country") + 
  labs(x="", y="Alpha diversity") +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12), axis.text.y = element_text(size = 12), axis.title = element_text(size = 14, face = 'bold')) +
  stat_compare_means(size = 4, label = "p.format", comparisons = country_comps)

group_comps <- list(c(1, 2), c(2, 3), c(1, 3))
ggplot(all_three_shannon_alpha, aes(x=Group, y=alpha)) + 
  ggtitle("By group") + 
  labs(x="", y="Alpha diversity") +
  geom_boxplot() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 12), axis.text.y = element_text(size = 12), axis.title = element_text(size = 14, face = 'bold')) +
  stat_compare_means(size = 4, label = "p.format", comparisons = group_comps)
  #geom_point(alpha = 0.3, position = "jitter")

country_group_comps <- list(c('Acute typhoid', 'Carrier'), c('Acute typhoid', 'Healthy Control'), c('Healthy Control', 'Carrier'))

ggboxplot(all_three_shannon_alpha, facet.by = "Country", y = "alpha", x = "Group", color = "Group") + stat_compare_means(comparisons = country_group_comps, label = 'p.signif') + rremove("x.text") + rremove("xlab") + rremove("x.ticks") # +rotate_x_text(angle = 45)


ggboxplot(all_three_shannon_alpha, facet.by = "Group", y = "alpha", x = "Country", color = "Country") + stat_compare_means(comparisons = country_comps) + rremove("x.text") + rremove("xlab") + rremove("x.ticks") # +rotate_x_text(angle = 45)


antibiotic_comps <- list(c('Yes', 'No'))
ggboxplot(all_three_shannon_alpha, facet.by = "Country", y = "alpha", x = "Antibiotics_taken_before_sampling_yes_no_assumptions", color = "Antibiotics_taken_before_sampling_yes_no_assumptions", legend.title = 'Antibiotics') + stat_compare_means(comparisons = antibiotic_comps) + rremove("x.text") + rremove("xlab") + rremove("x.ticks")


```

### blantyre

ANOVA

```{r}

anova_blantyre_sgaa_res <- read_delim(file.path(output_folder_blantyre, '3_alpha', 'shannon.anova.results.tsv'), delim = '\t')
anova_blantyre_sgaa_res %>% kbl %>% kable_styling()

```

Plots

- The p-value on the graph is a kruskal wallis test (which is just wilcoxon i think?)

```{r}
blantyre_shannon_alpha <- read_delim(file.path(output_folder_blantyre, '3_alpha', 'shannon.alpha_results.tsv'))
blantyre_shannon_alpha <- left_join(blantyre_shannon_alpha, meta_for_alpha, by = 'isolate')

my_comparisons <- list(c('No', 'Yes'), c('No', 'Yes'), c('No', 'Yes'))

ggplot(blantyre_shannon_alpha, aes(x=age_bracket, y=alpha, fill = Antibiotics_taken_before_sampling_yes_no_assumptions)) + 
  #ggtitle("All three countries - country by group") + 
  labs(x="Age bracket", y="Alpha diversity", fill='Antibiotics', color = 'Antibiotics') +
  geom_boxplot(outlier.shape = NA, alpha = 0.2) +
  #geom_point(aes(color=as.factor(Antibiotics_taken_before_sampling_yes_no_assumptions)), alpha = 0.3) +
  geom_point(alpha = 0.5, position=position_dodge(width=0.75),aes(group=Antibiotics_taken_before_sampling_yes_no_assumptions, color=as.factor(Antibiotics_taken_before_sampling_yes_no_assumptions))) +
  stat_compare_means(label.y = c(5.2, 5.5, 5.7), size = 4, label = "p.format")
  

```


### dhaka

```{r}

anova_dhaka_sgaa_res <- read_delim(file.path(output_folder_dhaka, '3_alpha', 'shannon.anova.results.tsv'), delim = '\t')
anova_dhaka_sgaa_res %>% kbl %>% kable_styling()

```

### kathmandu

```{r}

anova_kathmandu_sgaa_res <- read_delim(file.path(output_folder_kathmandu, '3_alpha', 'shannon.anova.results.tsv'), delim = '\t')
anova_kathmandu_sgaa_res %>% kbl %>% kable_styling()

```


## Beta diversity

Information on interpretation of PERMANOVA R2 from here - https://forum.qiime2.org/t/group-significance-changes-and-interpretation/2396 , some info on the p-values here https://forum.qiime2.org/t/understanding-beta-group-significance-permanova-results/12648

```{r}
run_plot_beta <- function(output_folder, meta, vars_to_plot){
  pcoa.data.handle <- file.path(output_folder, '4_beta', "first_2d_coords.txt")
  pcoa.var.handle <- file.path(output_folder, '4_beta', "pcoa_var.txt")
  pcoa.data <- read_delim(pcoa.data.handle, delim = " ", escape_double = FALSE, trim_ws = TRUE)
  #pcoa.data <- pcoa.data %>% separate(col = Y, sep = '\t', into = c('x', 'y')) %>% rename(ID = X)
  pcoa.var <- read_delim(pcoa.var.handle, delim = " ", escape_double = FALSE, trim_ws = TRUE)
  #pcoa.data <- read_tsv(pcoa.data.handle)
  
  meta_for_beta <- meta %>% select(isolate, Country, Group, Sex, age_bracket, Antibiotics_taken_before_sampling_yes_no_assumptions, group_country, group_antibiotic)
  
  pcoa.data <- left_join(pcoa.data, meta_for_beta, by = c('Sample' = 'isolate'))

  # add some different plots to the below
  # or, change this function to take the variable to colour by
  plots <- plot_beta(pcoa.data, pcoa.var, vars_to_plot)
  return(plots)
}
```

### All three countries

#### Plots

```{r}
all_three_vars_to_plot <- c('Country', 'Group', 'Sex', 'age_bracket', 'group_country', 'group_antibiotic')
all_three_plots <- run_plot_beta(output_folder_all_three, metadata, all_three_vars_to_plot)

a3_c <- all_three_plots$country_plot
a3_c
a3_g <- all_three_plots$group_plot
a3_g
all_three_plots$sex_plot
all_three_plots$age_plot
all_three_plots$group_country_plot
all_three_plots$group_antibiotic_plot


```

```{r fig.height = 3, fig.width = 3}
a3_c / a3_g
```


#### PERMANOVA

```{r}
pn_all3_csgaa_res <- read_delim(file.path(output_folder_all_three, '4_beta', 'permanova.results.tsv'))
pn_all3_csgaa_res %>% kbl %>% kable_styling()
```


### Just Blantyre

#### Plots

```{r}
blantyre_vars_to_plot <- c('Group', 'Sex', 'age_bracket', 'group_antibiotic')
blantyre_plots <- run_plot_beta(output_folder_blantyre, metadata, all_three_vars_to_plot)
b_g <- blantyre_plots$group_plot
b_g
blantyre_plots$sex_plot
blantyre_plots$age_plot
blantyre_plots$group_antibiotic_plot
```

#### PERMANOVA

```{r}
pn_blantyre_sgaa_res <- read_delim(file.path(output_folder_blantyre, '4_beta', 'permanova.results.tsv'))
pn_blantyre_sgaa_res %>% kbl %>% kable_styling()

```

### Just Dhaka

#### Plots

```{r}
dhaka_vars_to_plot <- c('Group', 'Sex', 'age_bracket', 'group_antibiotic')
dhaka_plots <- run_plot_beta(output_folder_dhaka, metadata, all_three_vars_to_plot)
d_g <- dhaka_plots$group_plot
d_g
dhaka_plots$sex_plot
dhaka_plots$age_plot
dhaka_plots$group_antibiotic_plot
```

#### PERMANOVA

```{r}
pn_dhaka_sgaa_res <- read_delim(file.path(output_folder_dhaka, '4_beta', 'permanova.results.tsv'))
pn_dhaka_sgaa_res %>% kbl %>% kable_styling()
```


### Just Kathmandu

#### Plots

```{r}
kathmandu_vars_to_plot <- c('Group', 'Sex', 'age_bracket', 'group_antibiotic')
kathmandu_plots <- run_plot_beta(output_folder_kathmandu, metadata, all_three_vars_to_plot)
k_g <- kathmandu_plots$group_plot
k_g
kathmandu_plots$sex_plot
kathmandu_plots$age_plot
kathmandu_plots$group_antibiotic_plot
```

#### PERMANOVA

```{r}
pn_kathmandu_sgaa_res <- read_delim(file.path(output_folder_kathmandu, '4_beta', 'permanova.results.tsv'))
pn_kathmandu_sgaa_res %>% kbl %>% kable_styling()
```

### Blantyre and Dhaka


#### Plots

```{r}
blantyre_dhaka_vars_to_plot <- c('Group', 'Sex', 'age_bracket', 'group_antibiotic')
blantyre_dhaka_plots <- run_plot_beta(output_folder_blantyre_dhaka, metadata, all_three_vars_to_plot)
blantyre_dhaka_plots$group_plot
blantyre_dhaka_plots$sex_plot
blantyre_dhaka_plots$age_plot
blantyre_dhaka_plots$group_antibiotic_plot
```

#### PERMANOVA

```{r}
pn_blantyre_dhaka_csgaa_res <- read_delim(file.path(output_folder_blantyre_dhaka, '4_beta', 'permanova.results.tsv'))
pn_blantyre_dhaka_csgaa_res %>% kbl %>% kable_styling()
```

### Combined plots

It looks like some might be being trimmed by the coordinate limits in the plot_beta function, but they're not (i've checked both blantyre and dhaka).

```{r fig.height= 3, fig.width = 3}
#b_g / d_g | k_g / plot_spacer() + plot_layout(guides = 'collect')
b_g + d_g + k_g + guide_area() + plot_layout(ncol = 2, guides = 'collect')
```

