
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message=FALSE)
```

## Sources
```{r imports}
library(kableExtra)
library(dplyr)
library(stringr)
source("/Users/flashton/Dropbox/GordonGroup/STRATAA_Microbiome/from_Leo/Leonardos_analysis/bin/core_functions.R")
source("/Users/flashton/Dropbox/GordonGroup/STRATAA_Microbiome/from_Leo/Leonardos_analysis/bin/config.R")
```

## paths

```{r paths}
strataa_disease_health_bigmap_data_path <- "~/Dropbox/GordonGroup/STRATAA_Microbiome/big-map/2023.05.01/analyse_output/tsv-results/all_RPKMs.transpose.csv"
strataa_carrier_health_bigmap_data_path <- "~/Dropbox/GordonGroup/STRATAA_Microbiome/big-map/2023.05.12/analyse_output_healthy_carrier/tsv-results/all_RPKMs.transpose.csv"

#strataa_bigmap_metadata_path <- "~/Dropbox/GordonGroup/STRATAA_Microbiome/big-map/2023.05.01/2023.05.01.big_map_input.maaslin.csv"
```

## Setting general params

```{r filename regex}
filename_regex = "\\d+_\\d+_\\d+" #for patch and strata
```

## read in metadata

```{r read_in_metadata}
metadata <- read_metadata(metadata_handle)
# View(metadata)

```


## Run MAASLIN2 taxonomic

```{r read_in_metaphlan, include=FALSE}

strataa_metaphlan_data = read.csv(file = file.path(metaphlan_input_folder, '2023.05.11.all_strataa_metaphlan.csv'), header= TRUE, sep = ",", row.names = 1, stringsAsFactors = FALSE, check.names=FALSE)
strataa_metaphlan_metadata = read.csv(file = file.path(metaphlan_input_folder, '2023.05.11.strataa_metadata.metaphlan.csv'), header = TRUE, sep = ",", row.names = 1, stringsAsFactors = FALSE)

```

```{r run_maaslin_metaphlan_healthy_typhoid include=FALSE}

#variables_for_analysis <- c("Group")
groups_to_analyse <- c('Acute_Typhi', 'Control_HealthySerosurvey')

variables_for_analysis <- c("Group", "Sex", "Age", "Antibiotics_taken_before_sampling_yes_no_assumptions")
run_maaslin(strataa_metaphlan_data, strataa_metaphlan_metadata, maaslin_taxonomic_output_root_folder, 'Bangladesh', groups_to_analyse, variables_for_analysis, 'NONE', 'LOG', NA, 'metaphlan')

variables_for_analysis <- c("Group", "Sex", "Age", "Antibiotics_taken_before_sampling_yes_no_assumptions", "sequencing_lane")

run_maaslin(strataa_metaphlan_data, strataa_metaphlan_metadata, maaslin_taxonomic_output_root_folder, 'Malawi', groups_to_analyse, variables_for_analysis, 'NONE', 'LOG', c("sequencing_lane,32580_7"), 'metaphlan')

variables_for_analysis <- c("Group", "Sex", "Age", "Antibiotics_taken_before_sampling_yes_no_assumptions")
# variables_for_analysis <- c("Group", "Sex")
run_maaslin(strataa_metaphlan_data, strataa_metaphlan_metadata, maaslin_taxonomic_output_root_folder, 'Nepal', groups_to_analyse, variables_for_analysis, 'NONE', 'LOG',  NA, 'metaphlan')
```


For healthy vs carrier, not including antibiotics becayse they're all "No"



```{r run_maaslin_metaphlan_healthy_carrier}

groups_to_analyse <- c('Carrier', 'Control_HealthySerosurvey')
#bgd_acute_healthy_output_dir <- file.path(maaslin_taxonomic_output_root_folder, paste())
variables_for_analysis <- c("Group", "Sex", "Age")
run_maaslin(strataa_metaphlan_data, strataa_metaphlan_metadata, maaslin_taxonomic_output_root_folder, 'Bangladesh', groups_to_analyse, variables_for_analysis, 'NONE', 'LOG', NA, 'metaphlan')

#variables_for_analysis <- c("Group", "Sex", "Age", "Antibiotics_taken_before_sampling_yes_no_assumptions", "sequencing_lane")
variables_for_analysis <- c("Group", "Sex", "Age", "sequencing_lane")
run_maaslin(strataa_metaphlan_data, strataa_metaphlan_metadata, maaslin_taxonomic_output_root_folder, 'Malawi', groups_to_analyse, variables_for_analysis, 'NONE', 'LOG',  c("sequencing_lane,32580_7"), 'metaphlan')

variables_for_analysis <- c("Group", "Sex", "Age")
run_maaslin(strataa_metaphlan_data, strataa_metaphlan_metadata, maaslin_taxonomic_output_root_folder, 'Nepal', groups_to_analyse, variables_for_analysis, 'NONE', 'LOG', NA, 'metaphlan')
```

health vs acute vs carrier

```{r}

groups_to_analyse <- c('Carrier', 'Control_HealthySerosurvey', 'Acute_Typhi')
variables_for_analysis <- c("Group", "Sex", "Age", "Antibiotics_taken_before_sampling_yes_no_assumptions", "sequencing_lane")
run_maaslin(strataa_metaphlan_data, strataa_metaphlan_metadata, maaslin_taxonomic_output_root_folder, 'Malawi', groups_to_analyse, variables_for_analysis, 'NONE', 'LOG', c("Group,Control_HealthySerosurvey", "Lane,32712_2"), 'metaphlan')


variables_for_analysis <- c("Group", "Sex", "Age", "Antibiotics_taken_before_sampling_yes_no_assumptions")
run_maaslin(strataa_metaphlan_data, strataa_metaphlan_metadata, maaslin_taxonomic_output_root_folder, 'Nepal', groups_to_analyse, variables_for_analysis, 'NONE', 'LOG', c("Group,Control_HealthySerosurvey"), 'metaphlan')

```



## run maaslin2 functional

```{r read_in_bigmap_results echo = T, results = 'hide'}

# strataa_bigmap_data = read.csv(strataa_bigmap_data_path, header= TRUE, sep = ",", row.names = 1, stringsAsFactors = FALSE, check.names=FALSE)
strataa_disease_health_bigmap_data = read.csv(file = strataa_disease_health_bigmap_data_path, header= TRUE, sep = ",", row.names = 1, stringsAsFactors = FALSE, check.names=FALSE)
strataa_carrier_health_bigmap_data = read.csv(file = strataa_carrier_health_bigmap_data_path, header= TRUE, sep = ",", row.names = 1, stringsAsFactors = FALSE, check.names=FALSE)

rownames(strataa_disease_health_bigmap_data) <- rownames(strataa_disease_health_bigmap_data) %>% str_replace_all('-', '.') %>% str_replace_all('#', '.')
rownames(strataa_carrier_health_bigmap_data) <- rownames(strataa_carrier_health_bigmap_data) %>% str_replace_all('-', '.') %>% str_replace_all('#', '.')

# strataa_bigmap_metadata is the same as strataa_metaphlan_metadata
#strataa_bigmap_metadata = read.csv(file = strataa_bigmap_metadata_path, header = TRUE, sep = ",", row.names = 1, stringsAsFactors = FALSE)

#strataa_bigmap_metadata <- strataa_bigmap_metadata %>% filter(Gender != '#N/A')
#strataa_bigmap_metadata$Age <- as.numeric(as.character(strataa_bigmap_metadata$Age))

strataa_bigmap_metadata <- metadata %>% mutate(isolate_bigmap = str_replace_all(isolate, '_', '.'))
rownames(strataa_bigmap_metadata) <- strataa_bigmap_metadata$isolate_bigmap

```

typhi vs health

- todo = haven't changed the function for including the prevalence, probably just need to filter/select the meta-data or something?

```{r }

groups_to_analyse <- c('Acute_Typhi', 'Control_HealthySerosurvey')

variables_for_analysis <- c("Group", "Sex", "Age", "Antibiotics_taken_before_sampling_yes_no_assumptions")
run_maaslin(strataa_disease_health_bigmap_data, strataa_bigmap_metadata, maaslin_functional_output_root_folder, 'Bangladesh', groups_to_analyse, variables_for_analysis, 'NONE', 'LOG', NA, 'bigmap')
variables_for_analysis <- c("Group", "Sex", "Age", "Antibiotics_taken_before_sampling_yes_no_assumptions", 'sequencing_lane')
run_maaslin(strataa_disease_health_bigmap_data, strataa_bigmap_metadata, maaslin_functional_output_root_folder, 'Malawi', groups_to_analyse, variables_for_analysis, 'NONE', 'LOG', c("sequencing_lane,32580_7"), 'bigmap')

```

carrier vs health

```{r}

groups_to_analyse <- c('Carrier', 'Control_HealthySerosurvey')

variables_for_analysis <- c("Group", "Sex", "Age")
# variables_for_analysis <- c("Group")
run_maaslin(strataa_carrier_health_bigmap_data, strataa_bigmap_metadata, maaslin_functional_output_root_folder, 'Bangladesh', groups_to_analyse, variables_for_analysis, 'NONE', 'LOG', NA, 'bigmap')
variables_for_analysis <- c("Group", "Sex", "Age", "sequencing_lane")
run_maaslin(strataa_carrier_health_bigmap_data, strataa_bigmap_metadata, maaslin_functional_output_root_folder, 'Malawi', groups_to_analyse, variables_for_analysis, 'NONE', 'LOG', c("Lane,32712_2"), 'bigmap')

```

